Staged Abstract Interpreters

### 2019-09-04 paper/60_cases.tex
revision

One of the challenges of modern static analysis is that program
usually depends on large libraries programs
\cite{toman_et_al:LIPIcs:2017:7121}. If we are able to analyze
programs and libraries separately and reuse the result without losing
precision, then we can reduce the overhead of repeatedly analyzing
libraries. Some static analyzers compute a summary for a function or a
module, which can be reused later. However, when lacking the context
information, those analyses can be too conservative or unsound, which
both lead to imprecision. In this section, we show that
meta-programming can provide a trade-off between summary-based modular
analysis and slow whole-program analysis, by utilizing an existing
whole-program analyzer.

As we have already seen from Section \ref{sai}, and recall that the
primitive operation @close@ can denote a $\lambda$ term to a next-stage Scala function.  The specializing of the staged abstract interpreter to
$\lambda$ terms is independent to other parts of the analyzed program.
\begin{lstlisting}
  def close(ev: Eval => Ans)(λ: Expr, ρ: Rep[Env]): Rep[Value]
\end{lstlisting}

Consider there is a library that contains a list map function @map@:
\begin{lstlisting}
  (define (map xs f) (if (null? xs) '() (cons (f (car xs)) (map (cdr xs) f))))
\end{lstlisting}

The specialization (0CFA and without store-widening) of @close@ on
@map@ generates the following next-stage Scala code, that is, the
function @x39@:
\begin{lstlisting}
  val x39 = {(x40: List[Set[AbsValue]],          // argument value list
              x41: Map[Addr, Set[AbsValue]],     // store
              x42: Map[Config, Set[ValSt]],      // in cache
              x43: Map[Config, Set[ValSt]]) =>   // out cache
    val x50 = List[Addr](ZCFAAddr("xs"), ZCFAAddr("f")) // addresses of arguments
    val x51 = x50.zip(x40)                              // addresses paired with their values
    val x61 = x51.foldLeft (x41) { case (x52, x53) =>   // prepare a new store that joins
      val x54 = x53._1;  val x55 = x53._2               //   the arguments and their values
        val x58 = x52.getOrElse(x54, x57)               //   into the latest store, initially x41
        val x59 = x58.union(x55);  x52 + (x54 -> x59) } 
    ... }
\end{lstlisting}

We borrow the notation from traditional modular analysis
\cite{DBLP:conf/cc/CousotC02}: such generated function is a
reusable artifact which can be used modularly, i.e., independent to its context. 
The generated functions already eliminate the interpretation overhead and
monadic overhead, therefore can be consider as partial summaries. Although, we
still need to run the generated function at next stage to obtain the analysis
result.  At next stage, we can provide different abstract arguments, store and
caches, depending on the context where it is called.

>>>>

One of the challenges of modern static analysis is that programs usually depend
on large libraries and analyzing these libraries is time consuming
\cite{toman_et_al:LIPIcs:2017:7121}.
If we are able to analyze programs and libraries separately and reuse the
result without losing precision, then we can reduce the overhead of repeatedly
analyzing libraries. Some static analyzers compute a summary for a function or
a module, which can be reused later. However, when lacking of the context
information, those analyses can be too conservative or unsound, which both lead
to imprecision. In this section, we show that meta-programming can provide a
trade-off between summary-based modular analysis and slow whole-program
analysis, by utilizing an existing whole-program analyzer.

As we have already seen from Section \ref{sai}, and recall that the primitive
operation @close@ can denote a $\lambda$ term to a next-stage Scala function.
The specializing of the staged abstract interpreter to a $\lambda$ term is
independent to other parts of the analyzed program.
\begin{lstlisting}
  def close(ev: Eval => Ans)(λ: Expr, ρ: Rep[Env]): Rep[Value]
\end{lstlisting}

We use Scheme as an example, and consider there is a library that contains a
list map function @map@:
\begin{lstlisting}
  (define (map xs f) (if (null? xs) '() (cons (f (car xs)) (map (cdr xs) f))))
\end{lstlisting}

The specialization (0CFA and without store-widening) of @close@ to
@map@ generates the following next-stage Scala code, that is, the
function @x39@:
\begin{lstlisting}
  val x39 = {(x40: List[Set[AbsValue]],          // argument value list
              x41: Map[Addr, Set[AbsValue]],     // store
              x42: Map[Config, Set[ValSt]],      // in cache
              x43: Map[Config, Set[ValSt]]) =>   // out cache
    val x50 = List[Addr](ZCFAAddr("xs"), ZCFAAddr("f")) // addresses of arguments
    val x51 = x50.zip(x40)                              // addresses paired with their values
    val x61 = x51.foldLeft (x41) { case (x52, x53) =>   // prepare a new store that joins
      val x54 = x53._1;  val x55 = x53._2               //   the arguments and their values
        val x58 = x52.getOrElse(x54, x57)               //   into the latest store, initially x41
        val x59 = x58.union(x55);  x52 + (x54 -> x59) } 
    ... }
\end{lstlisting}

We borrow the notation from traditional modular analysis
\cite{DBLP:conf/cc/CousotC02}: such generated function is a modularly reusable
artifact, i.e., independent to its context. 
The generated functions already eliminate the interpretation overhead and
monadic overhead, therefore can be consider as partial summaries. Although, we
still need to run the generated function at next stage to obtain the analysis
result.  At next stage, we can provide different abstract arguments, store and
caches, depending on the context where it is called.
